\Large \bf FlexAlloc: Dynamic Memory Partitioning\\ between the Corevisor and Hostvisor in SeKVM

### 2023-12-23 paper.tex
Filled in TODOs

We implement this fragmentation by manipulating the iterators used to traverse
the guest VM's stage-2 page table. In particular, the VM's stage-2 page table
is a four-level page table, comprising a page global directory (PGD), a
page upper directory (PUD), a page middle directory (PMD), and multiple page table
entries (PTEs). TODO (Matthew - Page table fragmentation)

@@

TODO (Matthew - DPM)

>>>>

We implement this fragmentation by manipulating the iterators that provide the next useable
page at a given level when traversing the stage-2 page table for the purpose of recording
an additional entry. In the HypSec implementation, each stage-2 page level's pages are
kept contiguous to one another and never intermixed with pages from other levels.
Each iterator starts at a particular offset into the VM's statically allocated stage-2
page table region, and increments each time a new page is consumed during a traversal
for use at that level. The design of these iterators means that they are limited to
only operating within a large, contiguous region.

The HypSec page global directory (PGD), page upper directory (PUD), and page middle directory
(PMD) iterators operate within a single contiguous 8MB region divided into varying sized
sub-regions for each level. The PGD iterator operates over a 16 * PAGE_SIZE region, the PUD
iterator over a SZ_2M - (16 * PAGE_SIZE) region, and the PMD iterator over a 3 * SZ_2M region.
Our implementation eliminates the need for all of these subregions to be contiguous to
one another and enables fragmentation within each subregion itself.

The FlexAlloc iterators, which can be thought of as page allocators specific to VM stage-2
page tables, operate over a pool of eight, non-contiguous 1MB regions. The FlexAlloc iterators
first exhaust all pages from a given 1MB region. Once the current region is exhausted, FlexAlloc
will jump the iterators to the next region in the pool. Subsequent traversals of the
stage-2 page tables with the alloc flag enabled will pull pages from the seperate,
non-contiguous region. By removing the requirement for contiguous 8MB regions, FlexAlloc
enables reliable dynamic allocation of stage-2 page table regions.

We implemented two versions of the FlexAlloc iterators. In the first version, we used a single,
generalized iterator shared across all levels of the stage-2 page tables. In the HypSec implementation,
the iterators are structured such that a PUD page will never be sandwiches between two PMD pages.
The first version of the FlexAlloc iterators ignores this semantic and instead is non-discrimiant
in regards to the particular level a page will be used in. We decided to stop development of this
version before it was fully functional in favor of version two which more closely resembles recent changes made to SeKVM's
stage-2 page table allocation scheme.

In the second version, we opt to enforce the semantic that pages at each specific level are
allocated from seperate pools and thus never mixed with one another. To accomplish this while still maintaining the same pool of non-contiguous 1MB regions,
we split the single, general iterator into nine distinct iterators. One iterator is used
for PUD pages, two iterators are used for PMD pages, and six iterators are used for PTE pages.
The number of iterators assigned to each level is based on the potential maximum number of
pages that could be consumed at that level along with how many distinct 1MB regions each level
can be allocated from. When a page is allocated at a particular level, the correct iterator is chosen based on
it's assigned level and whether or not its preceeding iterators have been exhausted.

The naming semantics of the iterators at each level can be confusing so we feel it is worth
clarifying here. The PGD iterator allocates pages to use as entries inside of a PGD page. This
implies that the PGD iterator, as an example, actually allocates pages that are used in the
subsequent level, which in this case is PUD. All iterators follow the pattern of allocating
pages for a given level to point to which are implicitly actually pages in the next level
of the stage-2 page tables.

@@

To enable dynamic allocation of EL2 memory in general and beyond just the scope of just
the stage-2 page tables, we proposed and designed, but did not complete the implementation for,
a dynamic page manager (DPM) interface. DPM would enable other "services" within EL2 to
utilize dynamic allocation for their memory regions without needing to handle the process
of receiving memory from the host.
